Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-05-06 Manuscript/manuscript.Rmd
methods edits (boundary prediction section)

We first established a *predicted TAD boundary area* (PTBA) as a collection of adjacent base pairs that have been identified as being a predicted TAD boundary with probability 1 according to the random forest model. We then implemented DBSCAN on the distance matrix (in base pairs) constructed from the midpoints of each PTBA using the R packacge *dbscan* (version 1.1-5) [@hahsler2019dbscan]. To identify the optimal number of clusters, *k*, the eps parameter ($\epsilon$) was assigned as the respective data resolution. The heuristic being that PTBA midpoints with genomic distances less than resolution should occupy the same designated cluster. The corresponding MinPts parameter was set to 3. We defined these *k* clusters of PTBAs as *predicted TAD boundary regions* (PTBR). [??? This deserves a schematic figure, before Figure 5. And part of it may be in Results] [??? I would switch the notation. First, you detect REGIONS. They can be adjacent, clustered together, forming AREA. Within this area, you can do pam clustering, and identify POINT]

@@

To precisely identify the PTBA most representative of each PTBR we performed an extension of partitioning around medoids (PAM), known as CLARA (Clustering Large Applications) for better scaling and more computational efficiency [@reynolds2006clustering; @wei2000empirical]. The midpoint of the corresponding cluster medoid, in this case a particular PTBA, was defined as a *predicted TAD boundary point* (PTBP). Thus, the hierarchical structure of our defined predicted TAD boundary objects were as follows: PTBR > PTBA > PTBP. That is, each PTBR is a collection of PTBA midpoints, with the uniquely identified PTBP being the most representative midpoint within a clustered PTBR.

>>>>

We first established a *predicted TAD boundary regions* (PTBR) as clusters of densitity-reachable base pairs which have been identified as being a predicted TAD boundary with probability 1 according to the random forest model using *DBSCAN* (version 1.1-5) [@hahsler2019dbscan]. The DBSCAN algorithm takes two parameters, *eps* ($\epsilon$) and *MinPts*. The eps parameter ($\epsilon$) was assigned as the respective data resolution. The heuristic being that base pairs with genomic distances less than resolution should occupy the same designated cluster. The corresponding MinPts parameter was set to 3. [??? This deserves a schematic figure, before Figure 5. And part of it may be in Results;!!!added PTBR to example of Figure 5 A] [??? I would switch the notation. First, you detect REGIONS. They can be adjacent, clustered together, forming AREA. Within this area, you can do pam clustering, and identify POINT;!!!editted, used term "region" instead of "area"]

To precisely identify the PTBA most representative base pair of each PTBR we performed an extension of partitioning around medoids (PAM), known as CLARA (Clustering Large Applications) for better scaling and more computational efficiency [@reynolds2006clustering; @wei2000empirical]. The midpoint of the corresponding cluster medoid, in this case a particular PTBR, was defined as a *preciseTAD boundary point* (PTBP). <!--Thus, the hierarchical structure of our defined predicted TAD boundary objects were as follows: PTBR > PTBA > PTBP.--> That is, each PTBR is a collection of base pairs, with the uniquely identified PTBP being the most representative base pair.

@@
